noam_data_exploration_v1

### 2024-08-20 noam_data_exploration_v3.Rmd
finished creating effects size plots for richness and slope models

```{r gls approach}

@@

slope_emm <- emmeans(slope_mod, ~ habitat*island_side) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ island_side | habitat)
# north and east coast not different from each other in backreef
# east and west not different from each other in fringing
# north and west not different from each other outer 10
# no sites different from each other outer 17
slope_est <- as.data.frame(slope_emm) # create df of emmeans estimates

@@

```{r glmm approach}
alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
# Framing - is adiv generally higher in certain habitats/regions?

@@

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  richness ~ year_factor*habitat*island_side + (1 | island_side/transect/location) + ar1(year_factor + 0 | location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
) # overspecified but going with it for now

summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
# emmeans(richness_mod_2, pairwise ~ habitat | year_factor) # if we want to compare habitats year to year
# model.sel(dredge(richness_mod_2))

>>>>

```{r gls approach, fig.height=10, fig.width=14}

@@

slope_emm <- emmeans(slope_mod, ~ habitat*island_side) # create emmeans object
emmeans::contrast(slope_emm, "consec", simple = "each", combine = TRUE, adjust = "bonferroni")
slope_plot_info <- plot(slope_emm, plotit = FALSE)
(emmeans_slope_graph <- 
  ggplot(data =slope_plot_info) +
  geom_point(aes(y = the.emmean, x = habitat, colour = island_side), size = 8, position = position_dodge(0.3)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL, x = habitat, colour = island_side), width = 0.2, size = 1.5,
                position = position_dodge(0.3) )+
  geom_hline(yintercept = 0) +
  theme_classic() +
  labs(y = "Estimated maringal mean slope\n", x = "Habitat") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )

# Note: not sure I want to include this
emmeans(slope_mod, pairwise ~ island_side | habitat)
# north and east coast not different from each other in backreef
# east and west not different from each other in fringing
# north and west not different from each other outer 10
# no sites different from each other outer 17
slope_est <- as.data.frame(slope_emm) # create df of emmeans estimates

@@

```{r glmm approach, fig.height = 20, fig.width=15}
alpha_diversity_quad$year_factor <- as.factor(alpha_diversity_quad$year) # for glmmTMB ar1 structure
# Framing - is adiv generally higher in certain habitats/regions?

@@

# probably the closest to what others have done:
richness_mod_2 <- glmmTMB(
  richness ~ year_factor*habitat*island_side + (1 | site/transect/location) + ar1(year_factor + 0 | location),
  family = poisson,
  data = alpha_diversity_quad,
  control = glmmTMBControl(optCtrl = list(iter.max = 2000, eval.max = 2000)), na.action = na.pass
) # overspecified but going with it for now

summary(richness_mod_2) # looks like strong autocorrelation, which makes sense to me
Anova(richness_mod_2)
plot(allEffects(richness_mod_2))
mod_emm <- emmeans(richness_mod_2, ~ habitat * island_side * year_factor) # if we want to compare habitats year to year
richness_plot_info <- plot(mod_emm, plotit = FALSE)
(emmeans_graph <- 
  ggplot(data =richness_plot_info) +
  geom_point(aes(x = exp(the.emmean), y = year_factor, colour = island_side), size = 8) +
  geom_errorbar(aes(xmin = exp(asymp.LCL), xmax = exp(asymp.UCL), y = year_factor, colour = island_side), width = 0.2, size = 1.5) +
  facet_wrap(~habitat) +
  theme_classic() +
  labs(y = "Year\n", x = "Estimated marginal mean richness") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
#        panel.background = element_blank(),
        legend.text = element_text(size=20),
        legend.title = element_text(size=20),
        axis.line = element_line(colour = "black"),
        axis.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = .5, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = 0, face = "plain"),
        strip.text.x = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain"),
        strip.text.y = element_text(color = "grey20", size = 25, hjust = .5, vjust = 0, face = "plain") ) )
